### 2018-01-12 manuscript.rmd
numbered sections, added abstract

## Introduction

@@

## Methods

@@

#### Collecting

@@

#### Sequencing

@@

#### Analysis

@@

## Results and Discussion

### Species phylogeny and hypothesis testing

@@

The broad sampling approach of this phylogeny provides new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae. This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focussed. Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, however there is weaker support for Clade B (Fig. 3A, 3B). Expanding sampling of species that probably fall in Clade B, including other *Erenna* species, rhodaliids, and relatives of Undescribed sp L, will greatly expand our understanding of these two groups and perhaps provide evidence of Clade B synapomorphies. Similarly, within Calycophorae, increased taxon sampling is needed. This study, and the previous study, suggest that the prayomorphs are paraphyletic, but for slightly different reasons. In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs. Expanded taxon sampling, particularly *P. dubia* or a nectopyramid, but also extensive sampling across the major prayomorph and diphyomorph groups, will expand our understanding of relationships within Calycophorae. This will be especially important for understanding trait evolution within Calycophorae,  for example, the release of eudoxids (Fig. 4), or the arrangement of male and female zooids along the stem (see below).

## Character Evolution
```{r char_evo_setup, include=F, echo=F}
#Load data
read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
rownames(cdata) = cdata$Species

@@

###Evolution of Monoecy

@@

###The Evolution of Zooid Types

@@

###The Gain and Loss of the Pneumatophore

@@

###The Gain and Loss of Tentilla

@@

###The Evolution of Vertical Habitat Use

@@

## Conclusions
Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans. We find support for two monophyletic groups within the clade that is sister to Pyrostephidae -- Calycophorae and a new clade we name Euphysonectae. Within Euphysonectae, we find support for two monophyletic groups dubbed Clade A and Clade B. Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores. These results are consistent with a hypothesis that monoecy evolved twice, once in Clade A and once in Calycophorae. This hypothesis is consistent with differences in the arrangement of male and female zooids in the two monoecious groups. Future studies with the addition of key taxa within Euphysonectae will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution, including the evolution of zooid types, that cannot be fully assessed here.

@@

This work was supported by the National Science Foundation (DEB-1256695 and the Waterman Award). Sequencing at the Brown Genomics Core facility was supported in part by NIH P30RR031153 and NSF EPSCoR EPS-1004057. Data transfer was supported by NSF RII-C2 EPS-1005789. Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island. SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster. We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph. We also thank the MBARI crews and ROV pilots for collection of the specimens.

>>>>

Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface of the ocean, like the familiar Portuguese man of war, to the deep sea. Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators in the open ocean. Previous work with a few molecular markers provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora. Here, we present a new siphonophore phylogeny, obtained by collecting new transcriptome data from 30 siphonophore species, and analyzing them in combination with 13 publicly available genomic and transcriptomic datasets. We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including gain and loss of zooid types, life history traits, and habitat. Sequence assembly, annotation, homology evaluation, and supermatrix construction were all conducted using the Agalma pipeline, and the supermatrix was subsequently analysed using Maximum Likelihood and Bayesian approaches. The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but finds strong support for new clades within Codonophora that were previously unresolved. These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose twice.

## 1. Introduction

@@

## 2. Material and methods

@@

#### 2.1 Collecting

@@

#### 2.2 Sequencing

@@

#### 2.3 Analysis

@@

## 3. Results and Discussion

### 3.1 Species phylogeny and hypothesis testing

@@

The broad sampling approach of this phylogeny provides new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae. This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focussed. Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, however there is weaker support for Clade B (Fig. 3A, 3B). Expanding sampling of species that probably fall in Clade B, including other *Erenna* species, rhodaliids, and relatives of Undescribed sp L, will greatly expand our understanding of these two groups and perhaps provide evidence of Clade B synapomorphies. Similarly, within Calycophorae, increased taxon sampling is needed. This study, and the previous study, suggest that the prayomorphs are paraphyletic, but for slightly different reasons. In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs. Expanded taxon sampling, particularly *P. dubia* or a nectopyramid, but also extensive sampling across the major prayomorph and diphyomorph groups, will expand our understanding of relationships within Calycophorae. This will be especially important for understanding trait evolution within Calycophorae,  for example, the release of eudoxids (Fig. 4), or the arrangement of male and female zooids along the stem (see section 3.2 below).

## 3.2 Character Evolution
```{r char_evo_setup, include=F, echo=F}
#Load data
read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
rownames(cdata) = cdata$Species

@@

### 3.3 Evolution of Monoecy

@@

### 3.4 The Evolution of Zooid Types

@@

### 3.5 The Gain and Loss of the Pneumatophore

@@

### 3.6 The Gain and Loss of Tentilla

@@

### 3.7 The Evolution of Vertical Habitat Use

@@

## 4. Conclusions
Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans. We find support for two monophyletic groups within the clade that is sister to Pyrostephidae -- Calycophorae and a new clade we name Euphysonectae. Within Euphysonectae, we find support for two monophyletic groups dubbed Clade A and Clade B. Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores. These results are consistent with a hypothesis that monoecy evolved twice, once in Clade A and once in Calycophorae. This hypothesis is consistent with differences in the arrangement of male and female zooids in the two monoecious groups. Future studies with the addition of key taxa within Euphysonectae will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution, including the evolution of zooid types, that cannot be fully assessed here.

@@

This work was supported by the National Science Foundation (DEB-1256695 and the Waterman Award). Sequencing at the Brown Genomics Core facility was supported in part by NIH P30RR031153 and NSF EPSCoR EPS-1004057. Data transfer was supported by NSF RII-C2 EPS-1005789. Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island. SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster. We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph. We also thank Zack Lewis for help with RNA extractions. We also thank the MBARI crews and ROV pilots for collection of the specimens.
